\nopagenumbers
{\bf Theorem 6.6}
\vskip 6pt
Let $G$ be a graph of order $n \geq 3$. If $$deg(u)+deg(v) \geq n$$
for each pair $u$, $v$ of non-adjacent vertices of $G$, then $G$ is {\it Hamiltonian}.

\vskip 10pt
{\bf Proof:}
\vskip 6pt
Assume, to the contrary, that there exists a non-Hamiltonian graph $G$ of order $n \geq 3$ such that
$deg(u)+deg(v) \geq n$ for each pair $u$, $v$ of non-adjacent vertices of $G$. It may be the case that if we
add certain edges to $G$, we obtain $K_n$, which is obviously Hamiltonian. Add as many edges as possible to $G$
so that the resulting graph $H$ is not Hamiltonian. Therefore, adding any edge to $H$ results in a Hamiltonian graph.
Also, $deg_H(u)+deg_H(v) \geq n$ for every pair of $u$, $v$ of non-adjacent vertices of $H$.
\vskip 1mm
Since $H$ is not complete, $H$ contains pairs of non-adjacent vertices. Let $x$ and $y$ be two non-adjacent
vertices of $H$. Thus $H+xy$ is Hamiltonian. Furthermore, every Hamiltonian cycle of $H+xy$ must contain the edge $xy$.
This means that $H$ contains a Hamiltonian $x$---$y$ path $P=(x=x_1,x_2, \ldots ,x_n=y)$. We claim that whenever 
$x_1x_i$ is an edge of $H$, where $2 \leq i \leq n$, then $x_{i-1}x_n$ is not an edge of $H$, for otherwise,
$$(x_1,x_i,x_{i+1}, \ldots , x_n,x_{i-1},x_{i-2}, \ldots ,x_1)$$
is a Hamiltonian cycle of $H$ is impossible. Hence for each vertex in $ \lbrace x_2,x_3, \ldots ,x_n \rbrace $ that is
adjacent to $x_1$, there is a vertex in $ \lbrace x_1,x_2, \ldots , x_{n-1} \rbrace $ that is not adjacent to $x_n$.
Howeve, this means that $deg(x_n) \leq (n-1)-deg(x_1) $ and so $$deg_H(x)+deg_H(y) \leq n-1$$ This is a contradiction.
\vskip 1in
{\bf Corollary 6.7}
\vskip 6pt
Let $G$ be a graph of order $n \geq 3$. {\it If} $deg(v) \geq {n \over 2}$ for each vertex $u$ of $G$, {\it then} $G$ is {\it Hamiltonian}.
\vskip 10pt
{\bf Proof:}
\vskip 6pt
Certainly, if $G=K_n$, then G is Hamiltonian. We may therefore assume that $G$ is not complete. Let $u$ and $v$ be two non-adjacent vertices of $G$.
Thus $$deg(u)+deg(v) \geq {n \over 2}+{n \over 2}=n$$ By Theorem 6.6, $G$ is Hamiltonian.

\vfill\eject
